The Processing of American Indian/Alaska Native Children through non-Tribal

### 2019-12-18 ms.rmd
update with proability models
ms update

```{r include=F}
rm(list=ls()); gc()
### for libra
setwd("~/Projects/ai_an_transitions")
library(tidyverse)
library(lubridate)
source("lifetable.r")

pop<-read_csv("./data/pop_aian_nat.csv")
afcars<-read_csv("./data/afcars_subset.csv",
                 col_types = cols(stfcid="c")) %>% 
  filter(age<18)

@@

ncands<-ncands %>% 
  mutate(year = year(rptdt)) %>% 
  filter(year>=2003)%>% 
  filter(chage<18)

### can't use FCID because of separate imputation

afcars %>% group_by(age, year) %>% summarise(n=n())
```

@@

this paper examines how american indian children are processed through us child welfare systems. it explores the likelihood that american indian children eventually end up in foster care, end up in foster care with non-indian foster parents, or see the rights of their parents terminated. I compare these pathways through the foster care system to the pathways experienced by children of other racial and ethnic groups.

## What am i estimating

@@

```{r ncandstables.pop}
ncands_first<-ncands %>% 
  group_by(chid) %>% 
  filter(rptdt==min(rptdt))

@@

investigation_tables<-bind_rows(investigation_tables)
```

@@

@@

```{r}
fc_inv<-make_conditionals(fc_tables, investigation_tables)
fc_sub<-make_conditionals(fc_tables, subst_tables)
```

@@

CONFIRM THAT THEIR IS ONLY ONE FCID PER CHID

temp<-ncands %>% 
  group_by(chid) %>% 
  summarise(n_fc=n_distinct(afcarsid))

>>>>

```{r include=F}
rm(list=ls()); gc()
library(tidyverse)
source("lifetable.r")

pop<-read_csv("./data/pop_aian_nat.csv")
afcars<-read_csv("./data/afcars_subset.csv",
                 col_types = cols(stfcid="c"))

afcars_all<-read_csv("./data/afcars_id_xwalk.csv",
                     col_types = cols(stfcid="c"))

@@

xwalk<-ncands_xwalk %>% 
  left_join(afcars_all) %>% 
  filter(!(is.na(fy))) %>% 
  select(stfcid, rptdt, fy) %>% 
  rename(fc_year = fy)

@@

This report examines how american indian children are processed through us child welfare systems. it explores the likelihood that american indian children eventually end up in foster care, end up in foster care with non-indian foster parents, or see the rights of their parents terminated. I compare these pathways through the foster care system to the pathways experienced by white and African American children.

I use the 2003 - 2016 AFCARS and NCANDS to calculate period lifetables for lifetime risk of investigation, substantiation, placement, placement into a non-ICWA compliant foster placement, and termination of parental rights. Each of these tables provides an estimate of the cumulative risk of experiencing the event before age 18 if risk-levels observed during the reported period remain constant.

## What am I estimating

@@

```{r ncandstables.pop}
ncands_first<-ncands %>% 
  select(.imp, chid, chage, rptdt) %>% 
  group_by(chid) %>% 
  filter(rptdt==min(rptdt)) %>% 
  distinct() %>% 
  ungroup() %>% 
  mutate(year = year(rptdt))

@@

investigation_tables<-bind_rows(investigation_tables)

investigation_tables %>% 
  filter(age==17) %>% 
  ggplot(aes(x=year, y = c)) +
  geom_line()

### way lower rates in 17
### check total case counts
### counts are low, probably some non-reporting states, removing for now
```

@@

## $P(entered, investigation)$ using ncands
```{r fctables.ncands}
ncands_fc_first<-ncands %>% 
  filter(ever_entered==T) %>% 
  group_by(stfcid) %>% 
  filter(chage==min(chage)) %>% 
  ungroup() %>% 
  select(stfcid, year, chage) %>% 
  distinct()

ncands_fc_nat<-ncands_fc_first %>% 
  group_by(year, chage) %>% 
  summarise(var = n()) %>% 
  rename(age = chage) %>% 
  left_join(pop) %>% 
  ungroup()

### run lifetable by year
ncands_fc_tables<-list()
years<-unique(ncands_fc_nat$year)
for(i in 1:length(years)){
  ncands_fc_tables[[i]]<-make_life_table(ncands_fc_nat %>% 
                  filter(year==years[i]))
}

ncands_fc_tables<-bind_rows(ncands_fc_tables)
```

## $P(investigation , entered)$ using afcars

```{r}
### among all entries, match on investigation
### question is afcars or ncands on left for join
### likely assymetry
### need first inv date in ncands_xwalk
### to ensure I have inv prior to fc
### to obtain P(inv,entered) where entered occurs after inv

afcars_inv<-afcars_first %>% 
  left_join(ncands_xwalk %>% 
              select(stfcid, rptdt) %>% 
              mutate(ncands_inv = T))

afcars_inv<-afcars_inv %>% 
  ungroup() %>% 
  mutate(rpt_year = year(rptdt)) %>% 
  filter(!(is.na(ncands_inv)))

## remove when year(rptdt)>year

afcars_inv<-afcars_inv %>% 
  filter(year>=rpt_year)

RECALL THAT

\[P(A|B) = \frac{P(A,B)}{P(B)}\]

ALSO, BAYES THEOREM

\[P(B|A) = \frac{P(A|B)P(B)}{P(A)}\]

for $A=FC$ and $B=investigation$, we wish to obtain $P(A|B)$. We obtain differing measurements of $P(A,B)$ from two datasets. I'll calculate each and compare.

@@

```{r}
fc_inv<-make_conditionals(fc_tables, investigation_tables)
fc_sub<-make_conditionals(fc_tables, subst_tables)
fc_ncands_inv<-make_conditionals(ncands_fc_tables, 
                                 investigation_tables)
fc_ncands_subst<-make_conditionals(ncands_fc_tables, 
                                 subst_tables)
```

NEXT: compare these to raw rates,

@@
